---
title: "GPT_prompting.Rdm"
author: "Fabienne Lind"
date: "April 2026"
output: html_document
---

Note: The following code was adapted by this blog post: https://rpubs.com/nirmal/setting_chat_gpt_R. 
and by the script shared as part of the supplementary materials of the following paper: 
Rathje, S., Mirea, D., Sucholutsky, I., Marjieh, R., Robertson, C., & Van Bavel, J. J. (2023, May 19). GPT is an effective tool for multilingual psychological text analysis. https://doi.org/10.31234/osf.io/sekf5
The script is explained in this video: https://www.youtube.com/watch?v=Mm3uoK4Fogc



### Install Required Packages

```{r}
library(httr)
library(tidyverse)
```


# Get API key

The first thing to to is to get a ChatGPT API key from here: https://platform.openai.com/overview 

Then, you could put your API key in the quotes below: 

```{r}
my_API <- "put_your_API_key_here"
```

The much saver option however (recommended!) is to save an API key in a save file. For example in an enviroment variable or in a file that is stored in a save folder. This way you do not share it directly in your script

How to do this: 
Option A: Environment variable

1. Set the environment variable. You need to do this only once. You can delete this line from your script later on.


```{r}
#Sys.setenv(GPT_API_KEY = "put_your_API_key_here")
```

2. Access the environment variable in your script. After storing the API key in your environment you can from now on call it with the following function.

```{r}
my_API <- Sys.getenv("GPT_API_KEY")
```

Option B: Excel file on your computer

```{r}
library(readxl)
setwd("/Users/fabienne/ucloud/Research/APIs")
api_keys <- read_excel("api_keys_recoded.xlsx")
my_API <- api_keys$GPT_API_KEY_courses

```


# GPT prompting

The "hey_chatGPT function will help you access the API and prompt GPT 


```{r}
library(httr)
library(jsonlite)
library(stringr)

# Make sure your API key is set
# my_API <- Sys.getenv("OPENAI_API_KEY")  # or set it directly (not recommended to hardcode)

hey_chatGPT <- function(answer_my_question, 
                        model = "gpt-4o-mini", 
                        temperature = 0, 
                        max_retries = 5, 
                        backoff_sec = 2) {
  url <- "https://api.openai.com/v1/chat/completions"
  
  for (attempt in 1:max_retries) {
    res <- tryCatch({
      POST(
        url = url,
        add_headers(Authorization = paste("Bearer", my_API)),
        content_type_json(),
        encode = "json",
        body = list(
          model = model,
          temperature = temperature,
          messages = list(
            list(role = "user", content = answer_my_question)
          )
        )
      )
    }, error = function(e) e)
    
    # Handle client-side errors
    if (inherits(res, "error")) {
      message(sprintf("Request error on attempt %d: %s", attempt, res$message))
      if (attempt < max_retries) Sys.sleep(backoff_sec * attempt) else return(character(0))
      next
    }
    
    # Check HTTP status
    if (http_error(res)) {
      # Try to parse error details
      parsed <- tryCatch(content(res, as = "parsed", type = "application/json"), error = function(e) NULL)
      if (!is.null(parsed) && !is.null(parsed$error) && !is.null(parsed$error$message)) {
        message(sprintf("API error (status %s) attempt %d: %s", status_code(res), attempt, parsed$error$message))
      } else {
        message(sprintf("HTTP error (status %s) on attempt %d", status_code(res), attempt))
      }
      if (attempt < max_retries) Sys.sleep(backoff_sec * attempt) else return(character(0))
      next
    }
    
    # Parse response JSON
    parsed <- tryCatch(content(res, as = "parsed", type = "application/json"), error = function(e) NULL)
    if (is.null(parsed) || is.null(parsed$choices) || length(parsed$choices) < 1 ||
        is.null(parsed$choices[[1]]$message$content)) {
      message(sprintf("Malformed response on attempt %d; retrying.", attempt))
      if (attempt < max_retries) Sys.sleep(backoff_sec * attempt) else return(character(0))
      next
    }
    
    answer <- str_trim(parsed$choices[[1]]$message$content)
    # Normalize: keep only 1 or 2; if model returns extra text, extract first digit 1/2
    digits <- str_extract_all(answer, "[12]")[[1]]
    if (length(digits) >= 1) {
      return(digits[1])
    } else {
      # If the model didn't follow instructions, treat as failure once and retry
      message(sprintf("Unexpected answer on attempt %d: '%s'", attempt, answer))
      if (attempt < max_retries) Sys.sleep(backoff_sec * attempt) else return(NA_character_)
    }
  }
  
  # If all retries exhausted
  return(NA_character_)
}



```



# Read in the dataset with the text column

We work here with the movie review data. 

This data was first used in Bo Pang and Lillian Lee, ``Seeing stars: Exploiting class relationships for sentiment categorization with respect to rating scales.'', Proceedings of the ACL, 2005. Please find more information on the dataset here.
https://huggingface.co/datasets/rotten_tomatoes


Using the OpenAPI costs something.
To not overburden the budget, we just test our code with a subset of the data. We use 20% of the data, the part for which we also predicted polarity with nb and svm.


```{r}

setwd("/Users/fabienne/ucloud/Lehre/Automated Text Analysis/PhD course text analysis/2024/data")
reviews_test <- read.csv("rotten_tomatoes_nb_svm_test.csv")

```

Let's briefly inspect the text column, named value here.

```{r}
head(reviews_test$text)
```

# Create a "gpt" column

We now add a new column to the dataframe. At the moment we just assign NAs. Out goal is now to fill this column with numbers that represent the polarity of the sentence.

```{r}
reviews_test$polarity_gpt <- NA
```

# Run a loop over your dataset and prompt ChatGPT 

The following loop includes a prompt for polarity. We allow two classes at the moment positive and negative. 

Note: In case you plan to use this for other data, replace `reviews_test` with the name of your dataframe and make sure that this line of code `text <- reviews_test[i,1]` assigns the column where your text is stored to the object text. 
```{r}

# Use only a small set
set.seed(42)  # optional, for reproducibility
idx <- sample(nrow(reviews_test), size = 10, replace = FALSE)
reviews_test_sample <- reviews_test[idx, , drop = FALSE]

```

```{r}

for (i in seq_len(nrow(reviews_test_sample))) {
  cat("Processing row", i, "\n")
  question <- "Is the sentiment of this text positive or negative? Answer only with a number: 1 if positive and 2 if negative. Here is the text:"
  text <- reviews_test_sample[i, 1]  # adjust column index/name if needed
  prompt <- paste(question, text)
  
  result <- hey_chatGPT(prompt)
  cat("Raw result:", result, "\n")
  reviews_test_sample$polarity_gpt[i] <- result
}


```



#Take only the first string from gpt and convert to a numeric 

```{r}
reviews_test$polarity_gpt <- substr(reviews_test$polarity_gpt, 1, 1)  
reviews_test$polarity_gpt <- as.numeric(reviews_test$polarity_gpt)

```

#Save the result

```{r}
setwd("/Users/fabienne/ucloud/Lehre/Automated Text Analysis/Summer School for Women Political Methodology Bremen/2026/data")
write.csv(reviews_test_sample, "rotten_tomatoes_gpt.csv", row.names = F)
reviews_test <- read.csv("rotten_tomatoes_gpt.csv")

```


#Comparison of manual coding decision with GPT proposal

We now inspect how well the manual annotations (stored in the column polarity) match up with the gpt annotations (column polarity_gpt).

#Recoding

First we need to bring the data into the same format.
We convert the polarity column to numbers

```{r}

reviews_test <- reviews_test %>%
    mutate(polarity = case_when(
      polarity == 'positive' ~ 1,
      polarity == 'negative' ~ 2
    ))

reviews_test <- reviews_test %>%
    mutate(polarity_nb = case_when(
      polarity_nb == 'positive' ~ 1,
      polarity_nb == 'negative' ~ 2
    ))

reviews_test <- reviews_test %>%
    mutate(polarity_svm = case_when(
      polarity_svm == 'positive' ~ 1,
      polarity_svm == 'negative' ~ 2
    ))

```


### Compute precision, recall and F1 per class.

```{r}

metrics <- function(reviews_test, categ) {
  results_table <- table(ifelse(reviews_test$polarity == categ, 1, 0), ifelse(reviews_test$polarity_gpt == categ, 1, 0))
  recall <- results_table[2, 2] / (results_table[2, 2] + results_table[2, 1])
  precision <- results_table[2, 2] / (results_table[2, 2] + results_table[1, 2])
  f1 <- 2 * recall * precision / (recall + precision)

  return(c(precision, recall, f1))
}

```


### Package all per-class metrics in one matrix

```{r}

package_metrics <- function(data, n_classes) { #n_classes is the number of categories here 2 (negative and positive)
  res <- matrix(NA, n_classes, 3) #3 relates to three metrics here precision, recall and F1
  for (c in 1:n_classes) {
    res[c, ] <- metrics(data, c)
  }
  return(res)
}

```

### Compute average F1 for the polarity analysis task

```{r}
macro_f1_sent <- function(reviews_test){
  res_pos <- metrics(reviews_test,1)# Metrics for positive class
  res_neg <- metrics(reviews_test,2)# Metrics for negative class
  avg_f1 <- mean(c(res_pos[3], res_neg[3])) ## Calculate average F1 score
  avg_f1
  return(avg_f1)
}
```

### Compute accuracy 

The diag() function extracts the diagonal elements of the contingency table, which correspond to the correctly predicted instances for each class.

The accuracy is then computed by dividing the sum of the correctly predicted instances by the total number of instances in the dataset (sum(res_table)).

```{r}
accuracy <- function(reviews_test){
  res_table <- table(reviews_test$polarity, reviews_test$polarity_gpt)
  acc <- sum(diag(res_table))/sum(res_table)
  return(acc)
}
```


### Run functions

Precision, recall, F1

```{r}
res_pos_sent <- metrics(reviews_test,1) # precision, recall, F1 for the 'positive' label in sentiment analysis
res_pos_sent

res_neg_sent <- metrics(reviews_test,2) # precision, recall; F1 for the 'negative' label in sentiment analysis
res_neg_sent
```



matrix with metrics for polarity analysis

```{r}
res_sent <- package_metrics(reviews_test, 2)
res_sent
```

average F1

```{r}
avg_f1 <- macro_f1_sent(reviews_test) 
avg_f1
```




