---
title: "2_lexicoder_sentiment_analysis.R"
output: html_document
date: "2025-07-08"
---


# Example Sentiment
https://quanteda.io/reference/data_dictionary_LSD2015.html
https://www.snsoroka.com/data-lexicoder/

Young, L. & Soroka, S. (2012). Affective News: The Automated Coding of Sentiment in Political Texts. doi:10.1080/10584609.2012.671234 . Political Communication, 29(2), 205–231

Important! Use the preprocessing script provided by Emily Luxon to improve the performance of the dictionary,
involving the detection of negations or handling of punctuation (download the script here: https://www.snsoroka.com/data-lexicoder/)


We go here with a toy corpus.

```{r}

library(quanteda) # text analysis project

# Create example corpus
texts_df <- data.frame(
  document = c("This aggressive policy will not win friends.",
           "The new initiative is not bad and might work well.")
)

texts_df$doc_id <- seq(1, nrow(texts_df)) # add id to dataframr
texts_df$doc_id <- as.character(texts_df$doc_id)

corp <- corpus(texts_df, text = "document", docid_field = "doc_id")
corp

```

Tokenize

```{r}

toks <- tokens(corp)
toks

```


Compound neg_negative and neg_positive tokens before creating a dfm object

```{r}
 
toks <- tokens_compound(toks, data_dictionary_LSD2015)
toks

```

Apply dictionary

```{r}

dfm_lookup(dfm(toks), data_dictionary_LSD2015)

#data_dictionary_LSD2015
```

Covert result to a dataframe and merge with 'texts_df'

```{r}
library(dplyr)
sentiment_dfm <- dfm_lookup(dfm(toks), data_dictionary_LSD2015)
sentiment_df <- convert(sentiment_dfm,to = "data.frame")

texts_sentiment_df <-  left_join(texts_df, sentiment_df, by = "doc_id")

```


Formula to calculate one sentiment score (normalized by text length): see 
score = (positive + neg_negative) - (negative + neg_positive) / total_words

Reference: 
Kouloumpis, E., Wilson, T. & Moore, J. (2011, July). Twitter sentiment analysis: The
good the bad and the omg!. In Proceedings of the Fifth International AAAI
conference on weblogs and social media, Barcelona, Spain (The AAAI press, pp. 538–
541).


```{r}

texts_sentiment_df$word_count <- lengths(toks)

texts_sentiment_df$score <- (
  texts_sentiment_df$positive +
  texts_sentiment_df$neg_negative -
  texts_sentiment_df$negative -
  texts_sentiment_df$neg_positive
) / texts_sentiment_df$word_count

```

### Other dictionaries in quanteda

List: 
https://rdrr.io/github/quanteda/quanteda.sentiment/f/vignettes/sentiment_analysis.Rmd




